---
title: "Do Scandals and Shocks Affect Elections?"
author: "Janet Hernandez"
date: "2022-10-24"
slug: "week7-shocks"
---

# Do Shocks Have a Significant Impact on Elections? #

This week, I will be doing the first blog extension. I will be replicating the Dobbs NYT example from section, but with the shock of recent inflation news. I predict that rising coverage on the state of the economy and inflation would have a negative affect on the President's party (Democrats) for generic ballot support. I am unsure whether or not this would also translate to an increase in support for Republicans but I am certain that I predict to see some hesitation towards Democrats due to economic inflation and market performances. I will be using the NYT developer API to source my data for news coverage. 

I will also try to modify and revise my existing predictive model from week 6. 

```{r setup, include=FALSE}
library(dotenv)
library(jsonlite)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
library(scales)
library(rjson)
library(jtools)
library(htmlwidgets)
library(reactable)
library(sqldf)
library(blogdown)
library(ggthemes)
library(plyr)

# load up hidden api key
article_api <- "3tPdNNFHIqqJySrXcFHjvgaFeCIac11M"
semantic_api <- Sys.getenv("SEMANTIC_API")
# 
# # set base url
base_url_art <- "http://api.nytimes.com/svc/search/v2/articlesearch.json?q="
base_url_sem <- "http://api.nytimes.com/svc/semantic/v2/concept/name"
# 
# # set parameters
term <- "inflation+biden"
facet_field <- "day_of_week"
facet <- "true"
fq <- "&fq=glocation='United States'"
begin_date <- "20220101"
end_date <- "20221022"


complete_url <-paste0(base_url_art,fq =term,"&facet_field=",facet_field,"&facet=",facet,"&begin_date=",begin_date,"&end_date=",end_date,"&api-key=",article_api,sep = "")
# import dataset to R

sus <- jsonlite::fromJSON(complete_url, flatten = TRUE)

# # view how many hits
sus$response$meta$hits


hits <- sus$response$meta$hits
cat("There were ",hits," hits for the search terms 'inflation' + 'biden' during 2022 to date",sep = "")
max_pages <- round((hits / 10) - 1)

# # store all pages in list
pages <- list()

#  trying again - WORKS!!!
Sys.sleep(1)
sus0 <-jsonlite::fromJSON(paste0(complete_url, "&page=0"), flatten = TRUE)
sus1 <- jsonlite::fromJSON(paste0(complete_url, "&page=1"), flatten = TRUE)
sus2 <- jsonlite::fromJSON(paste0(complete_url, "&page=2"), flatten = TRUE)
Sys.sleep(16)
sus3 <- jsonlite::fromJSON(paste0(complete_url, "&page=3"), flatten = TRUE)
sus4 <- jsonlite::fromJSON(paste0(complete_url, "&page=4"), flatten = TRUE)
Sys.sleep(16)
sus5 <- jsonlite::fromJSON(paste0(complete_url, "&page=5"), flatten = TRUE)
sus6 <- jsonlite::fromJSON(paste0(complete_url, "&page=6"), flatten = TRUE)
sus7 <- jsonlite::fromJSON(paste0(complete_url, "&page=7"), flatten = TRUE)
Sys.sleep(16)
sus8 <- jsonlite::fromJSON(paste0(complete_url, "&page=8"), flatten = TRUE)
sus9 <- jsonlite::fromJSON(paste0(complete_url, "&page=9"), flatten = TRUE)
Sys.sleep(12)
sus10 <- jsonlite::fromJSON(paste0(complete_url, "&page=10"), flatten = TRUE)
Sys.sleep(15)
sus11 <- jsonlite::fromJSON(paste0(complete_url, "&page=11"), flatten = TRUE)
sus12 <- jsonlite::fromJSON(paste0(complete_url, "&page=12"), flatten = TRUE)
Sys.sleep(12)
sus13 <- jsonlite::fromJSON(paste0(complete_url, "&page=13"), flatten = TRUE)
sus14 <- jsonlite::fromJSON(paste0(complete_url, "&page=14"), flatten = TRUE)
Sys.sleep(12)
sus15 <- jsonlite::fromJSON(paste0(complete_url, "&page=15"), flatten = TRUE)
Sys.sleep(5)




NYTSearch <- rbind_pages(list(sus0$response$docs, sus1$response$docs, sus2$response$docs,sus3$response$docs, sus4$response$docs, sus5$response$docs, sus6$response$docs, sus7$response$docs, sus8$response$docs, sus9$response$docs,sus10$response$docs, sus11$response$docs, sus12$response$docs, sus13$response$docs, sus14$response$docs, sus15$response$docs ))


# save df
saveRDS(NYTSearch, file = "inflation_2022.RDS")

# reload
mydata <- readRDS("inflation_2022.RDS")

```

## Background Literature on Shocks and Elections 

• Natural disaster: Shark attack (Achen and Bartels 2017), tornado
(Healy and Malhotra 2010) ⇝ decreased support for incumbents
• Sports (Healy, Mo, Malhotra 2010): college football team losing
their game ⇝ decreased support for incumbents
• Lottery (Bagues and Esteve-Volart 2016): Towns winning Spanish
Christmas lottery ⇝ increased support for incumbents
Global comparison (Powell & Whitten 1993): Punish only when
the gov’t is doing worse than other gov’ts
• Ex: the speed of recovery


```{r visualization, echo=FALSE}
# visualization by week
# extract raw date
mydata <- mydata %>% 
  mutate(publ_date = substr(pub_date, 1, 10))

# mutate week variable
mydata <- mydata %>% 
  mutate(week = strftime(publ_date, format = "%V"))
```

## New York Times Coverage on Biden and Inflation

Below is a plot of how the coverage of the terms 'inflation' + 'Biden' have changed from the period between January 2022 until October 15th, 2022. We can clearly see there is a huge spike in the months of August between weeks 31-33. I am looking at the first 15 pages of coverage from the NYT API. This equates to 160 articles out of a total of around 1000 written. 

In the next plot, I will compare how this coverage lines up with general trends in the general ballot support for Democrats and Republicans in the same time period. 

```{r, weekly plot, echo=FALSE}
# plot
mydata %>% 
  group_by(week) %>% 
  dplyr::summarize(count = n()) %>% 
  ggplot(aes(week, count, group = 1, color = count)) +
  geom_line() + theme_light()+ labs(y = "Article Count", x = "Week",
         title = "Weekly NYT Articles Mentioning 'Inflation' + 'Biden' in 2022",
         color = "")  

```


```{r, gen ballot setup, echo=FALSE, message=FALSE, warning=FALSE}
#now comparing this to generic ballot
X538_generic_ballot_averages_2018_2022 <- read.csv("~/Downloads/Section data/drive-download-20220922T150729Z-001/538_generic_ballot_averages_2018-2022.csv")

gb <- X538_generic_ballot_averages_2018_2022

# convert dat
gb <- gb %>%
  mutate(date_ = mdy(date)) %>%
  mutate(year = substr(date_, 1, 4)) %>%
  filter(year == 2022) %>%
  mutate(week = strftime(date_, format = "%V")) # Jan 1 looks weird 

#get avg by party and week
dem <- gb %>%
  filter(candidate == 'Democrats')
x <- plyr::ddply(dem, .(week), function(z) mean(z$pct_estimate))
x$candidate <- c('Democrats')
x$avg_dem <- x$V1
x <- x %>%
   select(-V1)
x$avg_dem <-  round(x$avg_dem , digits = 1)

rep <- gb %>%
   filter(candidate == 'Republicans')
y <- plyr::ddply(rep, .(week), function(z) mean(z$pct_estimate))
y$candidate <- c('Republicans')
y$avg_rep <- y$V1
y <- y %>%
  select(-V1)
y$avg_rep <-  round(y$avg_rep, digits = 1)
#
#put all data frames into list
df_list <- list(gb, x, y)
#
# #merge all data frames together
polls_df <- df_list %>% reduce(full_join, by=c("candidate", "week"))
#
# # remove NAs
polls_df[] <-  t(apply(polls_df, 1, function(x) c(x[!is.na(x)], x[is.na(x)])))
#
polls_df <- polls_df %>%
   select(-avg_rep)
#
polls_df$avg_support <- polls_df$avg_dem
#
polls_df <- polls_df %>%
  select(-avg_dem)
#
# # keep only unique dates
polls_df <- polls_df %>%
   distinct(cycle, week, date_, avg_support, candidate) %>%
   filter(week != 52)

# visualize polls
my_colors <- c("blue", "red")

```

## Peak Inflation Coverage Lines Up With Unexpected Drops in Support

```{r, plot for generic ballot, echo=FALSE, message=FALSE, warning=FALSE}
polls_df %>%
  group_by(candidate == 'Democrats') %>%
  mutate(date_ = as.Date(date_, format = '%Y-%m-%d')) %>%
  ggplot(aes(x = week, y = avg_support,
             colour = candidate)) +
  scale_color_manual(values = my_colors)+
   geom_line(aes(group=candidate), size = 0.3) + geom_point(size = 0.3) +
     #scale_x_date(date_labels = "%m, %Y") +
   ylab("Generic Ballot Support") + xlab("Week") +
      theme_light() + geom_segment(x=("31"), xend=("31"),y=0,yend=32, lty=2, color="black", alpha=0.3) +
      geom_segment(x=("33"), xend=("33"),y=0,yend=32, lty=2, color="black", alpha=0.3) + 
      annotate("text", x=("30"), y=27, label="Peak Inflation Coverage Window\n in NYT", size=4) 

```

Above is a graph that compares the peak weeks of coverage as evidenced in the plot above (Weeks 31-33) to the results of general ballot support pollings in the same time period. 

## Inflation and Biden News Punishing the Republicans More? 

An interesting observation of this graph in particular is how the dates line up perfectly with an overall drop in approval ratings for both Democrats and Republicans. However, the drop is much more significant for the non-incumbent party (Republicans) rather than President Biden's party. One possible explanation for the drop in both Democratic and Republican support could be that the coverage on inflation critiqued both sides, causing overall disillusion with the state of government and the economy. Interestingly enough, Republicans seem to be more punished than Democrats in this data. This may be explained or investigated further by creating sentiment scores for how the coverage was portraying Biden's response as either positive, negative, or neutral. Furthermore, the last thing to consider is that correlation does not equate to causation so there may be a possibility than other (possibly more important or popular) news coverage that I didn't consider in these weeks impacted Americans' support of the parties. 






