---
title: 'Week 6: The Ground Game'
author: "Janet Hernandez"
date: "2022-10-18"
slug: []
categories: []
tags: []
type: ''
subtitle: 'This week, I will focus on how the "Ground Game" or direct mobilization of voter turnout may affect election outcomes.'
image: ''
---
```{r, setup and lab code, include=FALSE}
options(blogdown.knit.on_save = TRUE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(collapse = TRUE)
options(warn = -1)
options(scipen = 999)

# read in district-level voting data
library(readr)
library(tidyverse)
library(ggplot2)
library(readr)
library(scales)
library(plotly)
library(rjson)
library(jtools)
library(htmlwidgets)
library(reactable)
library(sqldf)
library(blogdown)
require(knitr) # required for knitting from rmd to md
require(markdown) # required for md to html 

#Vote and seat share data from prev environments

voteseatshare_2020 <- read.csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house nationwide vote and seat share by party 1948-2020.csv")

house_party_vote_share_by_district_1948_2020 <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv")

## ECONOMIC DATA ##

# Data on GDP/quarter
economy <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/GDP_quarterly.csv")

# Data on RDI/quarter
RDI_quarterly <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/RDI_quarterly.csv")

# Data on unemployment
unemployment_national_quarterly_final <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/unemployment_national_quarterly_final.csv")

# Data on vote/seat share
popvote_df <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/house_popvote_seats.csv")

#state unemployment data
unemployment_state_monthly <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/unemployment_state_monthly.csv")


#DATA FROM FEC.gov
spending_bydistrict_2018 <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/ElectionBlog/2018_spending_by_district.csv")

house_party_vote_share_by_district <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv")%>%
filter(raceYear == 2018) %>%
select(State, raceYear, Area, RepCandidate, DemCandidate, RepVotesMajorPercent, DemVotesMajorPercent, st_fips, state_abb, CD, district_num, district_id, WinnerParty)

#Load data just for year 2018
house_vote_share_by_district_2018 <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv") %>%
filter(raceYear == 2018) %>%
select(State, raceYear, Area, RepCandidate, DemCandidate, RepVotesMajorPercent, DemVotesMajorPercent, st_fips, state_abb, CD, district_num, district_id, WinnerParty)

# Change from at large states w only one district to a code-able suffix
house_vote_share_by_district_2018$CD[grep(pattern = "-AL", x = house_vote_share_by_district_2018$CD)] <- c("AK-01","DE-01","MT-01","ND-01","SD-01", "VT-01","WY-01")

#Cleaned 2018 data from Ethan
ratings_share_2018 <-read.csv("~/Desktop/2018_ratings.csv")

#Incumbent list
incumbentslist <- read.csv("~/Desktop/incumb_dist_1948-2020 (3).csv")

# GROUND GAME LAB SECTION NOTES ##

dist_pv_df <- read.csv("~/Desktop/incumb_dist_1948-2020 (3).csv")
# read in cvap
cvap_district <- read.csv("~/Downloads/drive-download-20221016T234313Z-001/cvap_district_2012-2020_clean.csv")
# mutate geoid for merging
cvap_district <- cvap_district %>%
rename(st_cd_fips = geoid)

# select relevant years from voting data
table(dist_pv_df$year)
# 2012 - from 2018
# 2014, 2016, 2018, 2020 - from 2020
dist_pv_df <- dist_pv_df %>%
filter(year == 2018)
table(dist_pv_df$st_cd_fips)

# merge
dist_pv_cvap <- dist_pv_df %>%
inner_join(cvap_district, by = c('st_cd_fips', 'year'))

# mutate turnout
dist_pv_cvap <- dist_pv_cvap %>%
mutate(totalvotes = RepVotes + DemVotes,
turnout = totalvotes/cvap)

# mutate votes percent for glm
dist_pv_cvap <- dist_pv_cvap %>%
mutate(DemVotesMajorPct = DemVotesMajorPercent/100,
RepVotesMajorPct = RepVotesMajorPercent/100)

# drop uncontested seats
dist_pv_cvap_closed <- dist_pv_cvap %>%
filter(!is.na(DemCandidate), !is.na(RepCandidate)) %>%
mutate(DemVotesMajorPct = DemVotesMajorPercent/100,
RepVotesMajorPct = RepVotesMajorPercent/100)

```

```{r map data setup, include= FALSE}
require(sf)

# load geographic data
get_congress_map <- function(cong=114) {
tmp_file <- tempfile()
tmp_dir <- tempdir()
zp <- sprintf("https://cdmaps.polisci.ucla.edu/shp/districts114.zip",cong)
download.file(zp, tmp_file)
unzip(zipfile = tmp_file, exdir = tmp_dir)
fpath <- paste(tmp_dir, sprintf("districtShapes/districts114.shp",cong), sep = "/")
st_read(fpath)
}

# load 114th congress

cd114 <- get_congress_map(114)

cd114$DISTRICT <- as.numeric(cd114$DISTRICT)

cd114 <- cd114 %>% left_join(house_party_vote_share_by_district, by= c("STATENAME"= "State", "DISTRICT" = "district_num"))


#cd114 <- cd114 %>% inner_join(local_model_data, by= c("STATENAME"= "State", "DISTRICT" = "district_num"))

districts_simp <- rmapshaper::ms_simplify(cd114, keep = 0.01)
```


```{r theme janet, include = FALSE}
theme_janet <- function(base_size = 13, base_family = "") {
theme_grey(base_size = base_size, base_family = base_family) %+replace%
theme(

# Base elements which are not used directly but inherited by others
line =              element_line(colour = '#DADADA', size = 0.75,
linetype = 1, lineend = "butt"),
rect =              element_rect(fill = "#F0F0F0", colour = "#F0F0F0",
size = 0.5, linetype = 1),
text =              element_text(family = base_family, face = "plain",
colour = "#656565", size = base_size,
hjust = 0.5, vjust = 0.5, angle = 0,
lineheight = 0.9),

# Modified inheritance structure of text element
plot.title =        element_text(size = rel(1.5), family = '' ,
face = 'bold', hjust = -0.05,
vjust = 1.5, colour = '#3B3B3B'),
axis.title.x =      element_text(),
axis.title.y =      element_text(),
axis.text =         element_text(),

# Modified inheritance structure of line element
axis.ticks =        element_line(),
panel.grid.major =  element_line(),
panel.grid.minor =  element_blank(),

# Modified inheritance structure of rect element
plot.background =   element_rect(),
panel.background =  element_rect(),
legend.key =        element_rect(colour = '#DADADA'),

# Modifiying legend.position
legend.position = 'none',

complete = TRUE
)
}
```

# The Ground Game #

This week's focus is on mobilization of voters and how voter turnout may have an impact on who does well in elections. Most often, campaigns will focus on turning out those who are infrequent voters that align well with their party and or candidate's platform. Methods include texting, calling, and knocking on potential voter's doors to get out the vote. One big question I will try to investigate is whether or not there's a relationship between high turnout and the outcomes of either one of the major parties.

I will create a predictive model for turnout for 2022 and then test it against its predictions for 2018 versus actual outcomes to see how closely it will predict. Then, after I add in the variable of turnout into my existing model from previous weeks I will update my visualization and predictions.


## Visualizing Voter Turnout v. Party Major Vote Percent in 2018 ##

### Overall, Higher Turnout in 2018 Helped Democrats ###

```{r, rep plot for turnout, echo =FALSE}
# visualize
rep_turnout_plot <- ggplot(dist_pv_cvap_closed, aes(x = turnout, y = RepVotesMajorPercent)) +
geom_point() +
stat_smooth(method='lm', formula = y~x, color = "red")+
theme_janet() +
ggtitle("2018 Voter Turnout v.\n Republican Major Votes Percent")

ggplotly(rep_turnout_plot)

```

```{r, dem plot for turn out, echo=FALSE}
# visualize
dem_turnout_plot <- ggplot(dist_pv_cvap_closed, aes(x = turnout, y = DemVotesMajorPercent)) +
geom_point() +
stat_smooth(method='lm', formula = y~x, color= "blue")+
theme_janet() +
ggtitle("2018 Voter Turnout v.\n Democrat Major Votes Percent")

ggplotly(dem_turnout_plot)
```


Using just the 2018 data, I plotted the turnout and different parties major vote share percentage. As you can see, higher turnout had a slightly negative correlation to Republican Major Vote Percent while it had a positive correlation with Democratic Major Vote Percent. There are slightly more registered Democrats than Republicans in the US but Democrats are largely less active voters than Republican counterparts. This correlation in interesting to see at play. I think turnout will definitely help my model's fit this week.

**Note that 2018 results in comparison to the model in lab (which included more years) has a greater correlation it seems of voter turnout and democratic major vote pct. **


## Adding in Voter Turnout as a Predictor to Model ##

Model that uses previous midterm election data since 2012 to predict voter turnout for the 2022 election. I excluded presidential years since they skew data towards a higher turnout (Insert Citation) and 2022 doesn't happen to be a presidential year.

```{r adding turnout to datasets, echo=FALSE}
#Merge datasets from last week for local model and plot districts to see how it influences.

#Possible way to approach this is to take all the data i have from previous years on voter turnout on midterm years only and then create a table w a prediction for 2022 and append that to my last week model.

# read in cvap
turnout_gen <- read.csv("~/Downloads/drive-download-20221016T234313Z-001/cvap_district_2012-2020_clean.csv")
# mutate geo id for merging
turnout_gen <- turnout_gen %>%
rename(st_cd_fips = geoid)

turnout_gen <- incumbentslist %>%
inner_join(cvap_district, by = c('st_cd_fips', 'year'))

turnout_gen <- turnout_gen %>%
mutate(totalvotes = RepVotes + DemVotes,
turnout = totalvotes/cvap)
# Filter out for only midterm years and non-presidential bc I think that might have an influence / skew data to more turnout than we should expect empirically.
turnout_gen <- turnout_gen %>%
filter(year == 2012 | year == 2014 | year == 2018)

#Create linear function to then predict voter turnout.

turnout <- lm(turnout ~ year + state.x + president_party + cvap + winner_candidate_inc, data = turnout_gen)


summ(turnout)

turnoutpred_2022 <- data.frame()

turnoutpred <- as.data.frame(predict(turnout, new_data = turnoutpred_2022))

turnout_gen$prediction22 <- turnoutpred$`predict(turnout, new_data = turnoutpred_2022)`

#Added column for 2022 prediction now I can go ahead and map how well my model is prediction expected turnout v. actual turnout

#add margin column
turnout_gen <- turnout_gen %>%
mutate(margin = turnout - prediction22)

turnout_18 <- turnout_gen %>%
filter(year == 2018)

districts_simp <- districts_simp %>%
inner_join(turnout_18, by = 'district_id')

```

This model uses previous midterm election data since 2012 to predict voter turnout for the 2022 election. I excluded presidential years since they skew data towards a higher turnout (Insert Citation) and 2022 doesn't happen to be a presidential year. It has an R-squared of 0.69. It takes state, year, president_party, cvap, and incumbency into account. I will be using this to predict the overall Democratic Major Vote Percent as well as the seat distribution for the 2022 Midterm election.

```{r, local model data build, include=FALSE}
economy_q4 <- economy %>%
filter(quarter_yr == 4)

RDI_q4 <- RDI_quarterly %>%
filter(quarter_yr == 4)

unemployment_q4 <- unemployment_national_quarterly_final %>%
filter(quarter_yr == 4)

economy_model_data <- popvote_df %>%
inner_join(economy_q4, by = "year")

economy_model_data <-economy_model_data %>%
inner_join(RDI_q4, by = "year")
#Include District Level Economy Data
# Clean and update economy state level
unemployment_local <- unemployment_state_monthly %>%
filter(Year == 2018)

unemployment_local <- unemployment_local %>%
select('State.and.area', Year, Month, Unemployed_prct)

# Clean and include local RDI data for just 2018 and then join to economy data
RDI_local <- RDI_q4 %>%
filter(year == 2018)

# RDI and unemployment merged
unemployment_local <- unemployment_local %>%
inner_join(RDI_q4, by = c("Year" = "year"))
# Include GDP here too
economy_q4 <- economy_q4 %>%
select(year, GDP_growth_qt, GDP_growth_pct)

unemployment_local <- unemployment_local %>%
inner_join(economy_q4, by = c("Year" = "year"))

# Merge unemployment local to all state and district levels as equal. Merge by State.
unemployment_local <- rename(unemployment_local, State = 'State.and.area')


#Include District Level Polling Data (if possible)
#Using 2018 polling data to model our predictions
local_model_data_w6 <- house_vote_share_by_district_2018 %>%
inner_join(unemployment_local, by =  "State")


#Append the D and R Seats and national data to our local model for comparison in 2018.

natl_data_2018 <- economy_model_data %>%
select(year, R_seats, D_seats, R_majorvote_pct, D_majorvote_pct)

local_model_data_w6 <- local_model_data_w6 %>%
inner_join(natl_data_2018, by = c("Year" = "year"))

#Include District Incumbent Data

incumbentslist <- incumbentslist %>%
select(year, state, winner_party, district_id, winner_candidate_inc)

incumbentslist <- incumbentslist %>%
filter(year == 2018)

local_model_data_w6 <- local_model_data_w6 %>%
inner_join(incumbentslist, by = c("State" = "state","district_id" = "district_id"))

local_model_data_w6 <- local_model_data_w6 %>%
inner_join(ratings_share_2018, by = "CD")


## Ad spend

#Adding ad spending variable to our model

#Clean up ad data and add to local model then run lm

spending_bydistrict_2018 <- spending_bydistrict_2018 %>%
select(State, District, Candidate, Incumbent..Challenger.Open, Receipts)


#Need to add winning candidate name i think

local_model_data_w6 <- local_model_data_w6 %>%
mutate(winner_candidate = case_when( WinnerParty == "R" ~ local_model_data_w6$RepCandidate, WinnerParty == "D" ~ local_model_data_w6$DemCandidate))


local_model_data_w6 <- local_model_data_w6 %>%
inner_join(spending_bydistrict_2018, by = c("State" = "State", "district_num" ="District","winner_candidate_inc" = "Incumbent..Challenger.Open"))


local_model_data_w6$Receipts <- as.numeric(gsub("[\\$,]", "", local_model_data_w6$Receipts))
```

## Evaluating model's prediction for voter turnout vs. actual voter turnout in 2018. ##

Here, I am plotting differences in margin for turnout to test the accuracy of my prediction variable for 2022 turnout to add to my model later. Red indicates that the actual value is less than predicted value, therefore my model is under predicting in the red areas. The same goes for the blue. Where its more blue, such as in Florida, the predictive model is having a hard time and is over predicting these areas for voter turnout. I am curious to see what exact variables in certain states are causing this phenomenon.

```{r building w6 dataset, include=FALSE}
#Take prediction for 2022 turnout and add to my existing model
w6_df <- data.frame()

w6_df <-districts_simp %>%
left_join(local_model_data_w6, by = "district_id", "CD")

# drop na
w6_cleaned <- w6_df %>%
filter(!is.na(avg_margin))

#replace missing polling values with toss up rating
w6_df$avg <- w6_df$avg %>%
replace_na(4.7)

w6_df$Unemployed_prct <- w6_df$Unemployed_prct %>%
replace_na(3.6)

w6_df$Receipts <- w6_df$Receipts %>%
replace_na(960454)

#Simplify shapefile
w6_df <- rmapshaper::ms_simplify(w6_df, keep = 0.01)
```

```{r plot for margin between turnout , echo=FALSE}
turnoutmargin <- ggplot() +
geom_sf(data= w6_df,aes(fill= margin), color = "grey60", size = 0.05) +
scale_fill_gradient2(low = "red",
mid = "white",
high = "blue",
midpoint = 0,
name = "Difference in Margin") +
coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 52), expand = FALSE) +
theme_janet() +
theme(axis.title.x=element_blank(),
axis.text.x=element_blank(),
axis.ticks.x=element_blank(),
axis.title.y=element_blank(),
axis.text.y=element_blank(),
axis.ticks.y=element_blank(),
plot.title = element_text(margin = margin(0,0,10,0), hjust = 0.5)) +
labs(fill = "Voter Turnout", title = "Differences in Model Prediction v Actual Turnout in 2018\n Actual minus Predicted")

ggplotly(turnoutmargin)

```


# This Week's Model Prediction #

## Predicting for Major Democratic Vote Percent ##

```{r model for w6 dem vote pct, echo=FALSE}

w6_local_model <- lm(DemVotesMajorPercent.x ~ Unemployed_prct + winner_candidate_inc.x + Receipts + turnout + avg, data = w6_df)

summ(w6_local_model)

# add in Dem Vote prediction to w6_df and then filter for 2018, then append column to districts simp to plot
w6_df$DemVotePred22 <- predict(w6_local_model)

w6_simple <- w6_df %>%
select(State, Year, Month, cd, R_majorvote_pct, D_majorvote_pct, turnout, prediction22, DemVotesMajorPercent, DemVotePred22) %>%
filter(Year == 2018)

```
The R-squared of my model was slightly increased, from 0.74 to 0.75. Another important thing to note about this model is how significant the avg variable seems to be, which is indicative of the polling avg rating from 1 (Solid Democrat) to 7 (Solid Republican). The least important variable in terms of significance for this model is the rate of unemployment, suggesting that the economy is less of an indicator than the conditions of voter turnout, polling predictions, and money spent on ads.

## Visualizing 2022 Midterm Vote Prediction ##
```{r plot of model prediction2022, echo=FALSE}
w6_plot <- ggplot() +
geom_sf(data= w6_df, aes(fill= DemVotePred22), color = "grey60", size = 0.05) +
scale_fill_gradient2(low = "red",
mid = "white",
high = "blue",
midpoint = 50,
name = "Difference in Margin") +
coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 52), expand = FALSE) +
theme_janet() +
theme(axis.title.x=element_blank(),
axis.text.x=element_blank(),
axis.ticks.x=element_blank(),
axis.title.y=element_blank(),
axis.text.y=element_blank(),
axis.ticks.y=element_blank(),
plot.title = element_text(margin = margin(0,0,10,0), hjust = 0.5)) +
labs(fill = "DemMajorVotePct Prediction", title = "Week 6 Model Prediction\n DemMajorVotePct per District")

ggplotly(w6_plot)


#try to filter so that only one district from each state shows up 

districts22 <- w6_df %>% group_by(CD.x) %>% filter(row_number() == 1)

#summary(districts22$DemVotePred22 >= 50)

```


# Final Thoughts and Predictions #

My final prediction for week 6 is the following:

- **Mean of my Model's Democratic Major Vote Percent: 53%**

- **Model Prediction for Democratic Seats: 197 **

- **Model Prediction for Republican Seats: 231 **

## Notes from This Week / Issues ##

I tried to improve my model by doing a few things:

- Lost values when merging data sets and filled in NA's for districts within polling values to be set as 4.1667 for slightly leaning republican to account for the fact that the incumbent president's party usually does worse in midterms historically. This definitely helped make sense of my model and in my opinion delivered much better predictions, overshadowing some of the other variables I think. 

- Standardized any NA's in "ad spend data" with the mean of all.

- Added more years to my model in order to try to get a better picture and prediction rather than just one month of one year in previous models.

**Thoughts and Connections to Readings**

As noted by [Darr and Levendusky](https://hollis.harvard.edu/primo-explore/fulldisplay?docid=TN_cdi_gale_infotracacademiconefile_A375757032&context=PC&vid=HVD2&search_scope=everything&tab=everything&lang=en_US), campaign field offices and their prominence might be a better indicator of how the "Ground Game" is working. 

In [Enos and Fowler's](https://hollis.harvard.edu/primo-explore/fulldisplay?docid=TN_cdi_proquest_journals_2117059848&context=PC&vid=HVD2&search_scope=everything&tab=everything&lang=en_US) they  estimate that the 2012 presidential campaigns increased turnout in highly targeted states by 7–8 percentage points, on average, indicating that modern campaigns can significantly alter the size and composition of the voting population. Further evidence suggests that the predominant mechanism behind this effect is traditional ground campaigning, which has dramatically increased in scale in the last few presidential elections. What is unclear though for at least my purposes, is how has ground campaigning been affected in non-presidential election years? I estimate that this may be a overarching trend and that my model which tends to over predict slightly on voter turnout actually may be more helpful in accounting for these changes in historical data. 

Overall, one thing to note about my model (which I spoke to Professor Enos about) is how heavily it relies on polling data for its predictive power. I think this is very indicative of how polls can be good predictors on empirical grounds for how an election will sway. 

# References #

Joshua P Darr and Matthew S Levendusky. Relying on the Ground Game: The Placement and Effect of Campaign Field Offices. American Politics Research, 42(3):529–548, 2014. ISSN 1532-673X. URL: https://hollis.harvard.edu/primo-explore/fulldisplay?docid=TN_cdi_gale_infotracacademiconefile_A375757032&context=PC&vid=HVD2& search_scope=everything&tab=everything&lang=en_US.

Ryan D. Enos and Anthony Fowler. Aggregate Effects of Large-Scale Campaigns on Voter Turnout. Political Science Research and Methods, 6(4):733–751, 2016. URL https://hollis.harvard.edu/primo-explore/fulldisplay?docid=TN_cdi_ proquest_journals_2117059848&context=PC&vid=HVD2&search_scope=everything& tab=everything&lang=en_US.

Ryan D Enos and Eitan D Hersh. Party Activists as Campaign Advertisers: The Ground Campaign as a Principal-Agent Problem. American Political Science Review, 109(02): 252–278, 2015. URL: https://hollis.harvard.edu/primo-explore/fulldisplay?docid=TN_cdi_proquest_journals_1675033350&context=PC&vid=HVD2&search_ scope=everything&tab=everything&lang=en_US.






