---
title: 'Week 5: Ad Spend and Election Outcomes'
author: Janet Hernandez
date: '2022-10-08'
summary: This week, I look at how campaign ad spending has an effect on my existing
  models. I will specifically look at FEC election data and see if I can find a strong
  indicator of prediction for my model this week using ad data.
---
# Advertising and Election Outcomes # 

In this post, I will investigate the effects that ad spending can have on the outcome of elections on a district by district level. Using my current model, I will incorporate data found on the FEC's website to approximate ad spend on a campaign district level. Using how much money was spent will be my variable, though I am aware that some scholarship says that number of ads rather than money spent is also a good indicator. I will also run tests on that next week. I will also be using data from the 2018 midterm to build my predictive model. 

## Current Scholarship ##

Political scientists [Gregory A Huber and Kevin Arceneaux](https://hollis.harvard.edu/primo-explore/fulldisplay?docid=TN_cdi_proquest_miscellaneous_59786011&context=PC&vid=HVD2&search_scope=everything&tab=everything&lang=en_US) have noted in their research that most districts tend to be non-competitive. They also argue that "Advertising does a little to inform, next to nothing to mobilize, and a great deal to persuade potential voters." A key takeaway from their work is how campaign advertising can do a great deal to persuade potential voters to candidates. Therefore, for undecided voters, it is highly likely that receiving more of a certain candidates ads more often than another will increase the likelihood of that voter to vote for the candidate which they saw more ads for. 

This has pretty important implications and would be useful to include into our model and test whether or not the amount campaigns spend on ads and air space can affect our model's prediction and accuracy for both 2018 and the upcoming 2022 election. 
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(collapse = TRUE)
options(warn = -1)

library(tidyverse)
library(ggplot2)
library(readr)
library(scales)
library(plotly)
library(rjson)
library(jtools)
library(htmlwidgets)
library(reactable)
library(sqldf)


voteseatshare_2020 <- read.csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house nationwide vote and seat share by party 1948-2020.csv")

house_party_vote_share_by_district_1948_2020 <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv")


# Data on GDP/quarter
economy <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/GDP_quarterly.csv")

# Data on RDI/quarter
RDI_quarterly <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/RDI_quarterly.csv") 

# Data on unemployment
unemployment_national_quarterly_final <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/unemployment_national_quarterly_final.csv") 

# Data on vote/seat share
popvote_df <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/house_popvote_seats.csv")

#state unemployment data
unemployment_state_monthly <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/unemployment_state_monthly.csv")


#DATA FROM FEC.gov
spending_bydistrict_2018 <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/ElectionBlog/2018_spending_by_district.csv")

house_party_vote_share_by_district <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv")%>% 
  filter(raceYear == 2018) %>%
  select(State, raceYear, Area, RepCandidate, DemCandidate, RepVotesMajorPercent, DemVotesMajorPercent, st_fips, state_abb, CD, district_num, district_id, WinnerParty)

#Load data just for year 2018 
house_vote_share_by_district_2018 <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv") %>%
  filter(raceYear == 2018) %>%
  select(State, raceYear, Area, RepCandidate, DemCandidate, RepVotesMajorPercent, DemVotesMajorPercent, st_fips, state_abb, CD, district_num, district_id, WinnerParty)

# Change from at large states w only one district to a code-able suffix
house_vote_share_by_district_2018$CD[grep(pattern = "-AL", x = house_vote_share_by_district_2018$CD)] <- c("AK-01","DE-01","MT-01","ND-01","SD-01", "VT-01","WY-01")

#Cleaned 2018 data from Ethan  
ratings_share_2018 <-read.csv("~/Desktop/2018_ratings.csv")

#Incumbent list
incumbentslist <- read.csv("~/Desktop/incumb_dist_1948-2020 (3).csv")

```

```{r map data, include= FALSE}
require(sf)

# load geographic data
get_congress_map <- function(cong=114) {
tmp_file <- tempfile()
tmp_dir <- tempdir()
zp <- sprintf("https://cdmaps.polisci.ucla.edu/shp/districts114.zip",cong)
download.file(zp, tmp_file)
unzip(zipfile = tmp_file, exdir = tmp_dir)
fpath <- paste(tmp_dir, sprintf("districtShapes/districts114.shp",cong), sep = "/")
st_read(fpath)
}

# load 114th congress

cd114 <- get_congress_map(114)

cd114$DISTRICT <- as.numeric(cd114$DISTRICT)

cd114 <- cd114 %>% left_join(house_party_vote_share_by_district, by= c("STATENAME"= "State", "DISTRICT" = "district_num"))


#cd114 <- cd114 %>% inner_join(local_model_data, by= c("STATENAME"= "State", "DISTRICT" = "district_num"))

districts_simp <- rmapshaper::ms_simplify(cd114, keep = 0.01)
```

```{r clean national data, include= FALSE}
unemployment_national_quarterly_final <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/unemployment_national_quarterly_final.csv")

economy_q4 <- economy %>% 
  filter(quarter_yr == 4) 

RDI_q4 <- RDI_quarterly %>%
  filter(quarter_yr == 4)

unemployment_q4 <- unemployment_national_quarterly_final %>%
  filter(quarter_yr == 4)

economy_model_data <- popvote_df %>% 
  inner_join(economy_q4, by = "year")

economy_model_data <-economy_model_data %>% 
  inner_join(RDI_q4, by = "year")

national_model_data <-economy_model_data %>% 
  inner_join(unemployment_q4, by = "year")

```

# My Existing Model # 

My current model from Week 4 included incumbency, economic factors, and basic polling. Currently, the R squared for this model is at .73, a fair amount. 

```{r local model data, echo=FALSE}

#Building model off of 2018 data for simple purposes. 

#Include District Level Economy Data
  # Clean and update economy state level 
unemployment_local <- unemployment_state_monthly %>% 
  filter(Year == 2018, Month == 10)

unemployment_local <- unemployment_local %>% 
  select('State.and.area', Year, Month, Unemployed_prct)

# Clean and include local RDI data for just 2018 and then join to economy data
  RDI_local <- RDI_q4 %>%
    filter(year == 2018)
  
# RDI and unemployment merged
unemployment_local <- unemployment_local %>%
  inner_join(RDI_q4, by = c("Year" = "year"))
# Include GDP here too 
economy_q4 <- economy_q4 %>%
  select(year, GDP_growth_qt, GDP_growth_pct)

unemployment_local <- unemployment_local %>%
  inner_join(economy_q4, by = c("Year" = "year"))

# Merge unemployment local to all state and district levels as equal. Merge by State. 
unemployment_local <- rename(unemployment_local, State = 'State.and.area')


#Include District Level Polling Data (if possible)
  #Using 2018 polling data to model our predictions
local_model_data <- house_vote_share_by_district_2018 %>% 
  inner_join(unemployment_local, by =  "State")


#Append the D and R Seats and national data to our local model for comparison in 2018. 

natl_data_2018 <- economy_model_data %>% 
  select(year, R_seats, D_seats, R_majorvote_pct, D_majorvote_pct)

local_model_data <- local_model_data %>% 
  inner_join(natl_data_2018, by = c("Year" = "year"))

#Include District Incumbent Data 

incumbentslist <- incumbentslist %>% 
  select(year, state, winner_party, district_id, winner_candidate_inc)

incumbentslist <- incumbentslist %>% 
  filter(year == 2018)

#WHY IS IT ADDING 2 MORE VALUES IDK???!!! 
local_model_data <- local_model_data %>% 
  inner_join(incumbentslist, by = c("State" = "state","district_id" = "district_id"))

local_model_data <- local_model_data %>% 
  inner_join(ratings_share_2018, by = "CD")

w4_local_model <- lm(DemVotesMajorPercent ~ avg + Unemployed_prct + winner_candidate_inc, data = local_model_data)

summ(w4_local_model)

```

My new local model that includes district level data on polling, incumbency, and local employment data is much more accurate than before. However, the R-squared of 0.73 which is the highest so far only when I use DemMajorVotePct as my outcome variable. When I do DemSeats(which my previous models used) as my outcome variable, I get a lower R-squared of 0.50 exactly. 


```{r adding ad spend into model, include= FALSE}

#Adding ad spending variable to our model 

  #Clean up ad data and add to local model then run lm 

spending_bydistrict_2018 <- spending_bydistrict_2018 %>%
  select(State, District, Candidate, Incumbent..Challenger.Open, Receipts)


#Need to add winning candidate name i think

local_model_data <- local_model_data %>%
  mutate(winner_candidate = case_when( WinnerParty == "R" ~ local_model_data$RepCandidate, WinnerParty == "D" ~ local_model_data$DemCandidate))


local_model_data <- local_model_data %>% 
  inner_join(spending_bydistrict_2018, by = c("State" = "State", "district_num" ="District","winner_candidate_inc" = "Incumbent..Challenger.Open"))


local_model_data$Receipts <- as.numeric(gsub("[\\$,]", "", local_model_data$Receipts))

#cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
#cd114 <- cd114 %>% inner_join(local_model_data, by=c("DISTRICT"= "district_num", "STATENAME"= "State"))

```

```{r theme janet, include = FALSE}
theme_janet <- function(base_size = 13, base_family = "") {
  theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(
      
      # Base elements which are not used directly but inherited by others
      line =              element_line(colour = '#DADADA', size = 0.75, 
                                       linetype = 1, lineend = "butt"),
      rect =              element_rect(fill = "#F0F0F0", colour = "#F0F0F0", 
                                       size = 0.5, linetype = 1),
      text =              element_text(family = base_family, face = "plain",
                                       colour = "#656565", size = base_size,
                                       hjust = 0.5, vjust = 0.5, angle = 0, 
                                       lineheight = 0.9),
      
      # Modified inheritance structure of text element
      plot.title =        element_text(size = rel(1.5), family = '' , 
                                       face = 'bold', hjust = -0.05, 
                                       vjust = 1.5, colour = '#3B3B3B'),
      axis.title.x =      element_text(),
      axis.title.y =      element_text(),
      axis.text =         element_text(),
      
      # Modified inheritance structure of line element
      axis.ticks =        element_line(),
      panel.grid.major =  element_line(),
      panel.grid.minor =  element_blank(),
      
      # Modified inheritance structure of rect element
      plot.background =   element_rect(),
      panel.background =  element_rect(),
      legend.key =        element_rect(colour = '#DADADA'),
      
      # Modifiying legend.position
      legend.position = 'none',
      
      complete = TRUE
    )
}
```

Below is a plot of the actual outcomes from the 2018 election from which im pulling my predictive modeling data from.

```{r actual plot, echo= FALSE}

Actual2022_w5 <- ggplot() + 
geom_sf(data= districts_simp,aes(fill= DemVotesMajorPercent), color = "grey60", size = 0.05) + 
 scale_fill_gradient2(low = "red", 
                       mid = "white", 
                       high = "blue", 
                       midpoint = 50,
                       name = "DemMajorVotePct") +
  coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 52), expand = FALSE) +  
  theme_janet() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.title = element_text(margin = margin(0,0,10,0), hjust = 0.5)) +
  labs(fill = "Dem Major Vote Pct", title = "Actual Dem Major Vote Percent in 2018")


ggplotly(Actual2022_w5)

```

Now I have to add the variable of ad spending on a local level. I've gone ahead and downloaded data from the FEC for 2018 election spending data. This isn't exactly the ad spend per campaign but I am using it as a proxy by making the assumption that the more money a particular race / candidate has overall translates to how much it is spending in ads.

```{r plots, echo= FALSE}
w5_local_model <- lm(DemVotesMajorPercent ~ avg + Unemployed_prct + winner_candidate_inc + Receipts, data = local_model_data)

districts_simp$prediction <-head(predict(w5_local_model), n= 436)

#Add prediction to dataset to plot district by district prediction

Prediction2022_w5 <- ggplot() + 
geom_sf(data= districts_simp,aes(fill= prediction), color = "grey60", size = 0.05) + 
 scale_fill_gradient2(low = "red", 
                       mid = "white", 
                       high = "blue", 
                       midpoint = 50,
                       name = "DemMajorVotePct") +
  coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 52), expand = FALSE) +  
  theme_janet() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.title = element_text(margin = margin(0,0,10,0), hjust = 0.5)) +
  labs(fill = "Dem Major Vote Pct Prediction", title = "Model Prediction for Dem Major Vote Pct in 2022")


ggplotly(Prediction2022_w5)
```


```{r predict v expected plot, echo= FALSE}
summ(w5_local_model)

w5_model <- data.frame()

data_mod <- data.frame( District = local_model_data$CD,
  Predicted = predict(w5_local_model, new_data = w5_model),  
                       Observed = local_model_data$DemVotesMajorPercent)


Model <- ggplot(data_mod,                                  
       aes(x = Predicted,
           y = Observed))  +
  geom_point() + 
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              size = 2) +
 geom_label(label= data_mod$District) + theme_janet() + ggtitle("Predictive Model 3:\nPredicted v. Observed Democrat Major Vote Percent") 


ggplotly(Model)

```

# References #

Bafumi, J., Erikson, R., & Wlezien, C. (2018). Forecasting the 2018 Midterm Election using National Polls and District Information. PS: Political Science & Politics, 51(S1), 7-11. doi:10.1017/S1049096518001579

Ballotpedia. (2018). United States House of Representatives elections, 2018. https://ballotpedia.org/United_States_House_of_Representatives_elections,_2018

Ballotpedia. (2022). United States Congress elections, 2022. https://ballotpedia.org/United_States_Congress_elections,_2022

Congressional candidate data summary tables - FEC.gov. (2022). Retrieved 16 October 2022, from https://www.fec.gov/campaign-finance-data/congressional-candidate-data-summary-tables/?year=2022&segment=18

Cook Political Report. (2022). PVI Map and District List. https://www.cookpolitical.com/cook-pvi/2022-partisan-voting-index/district-map-and-list

Gerber, A.S., Gimpel, J. G., Green, D. P., & Shaw, D. R. (2011). How Large and Long-lasting Are the Persuasive Effects of Televised Campaign Ads? Results from a Randomized Field Experiment. American Political Science Review, 105(1), 135–150. https://doi.org/10.1017/S000305541000047X

Wesleyan Media Project. (2022, October 6). Democrats Out-Pacing GOP in Senate Races. https://mediaproject.wesleyan.edu/releases-100622/
