---
title: Post Election Reflection
author: Janet Hernandez
date: '2022-11-20'
slug: []
categories: []
tags: []
type: ''
subtitle: 'Post-election Reflection on Model - Gov 1347, Fall 2022'
image: ''
---
# Introduction #
As of Monday, November 21st, 2022, the current makeup of the 2022 Congress is that 430/435 seats have been called and currently have confirmed 218 Republican seats, 213 Democratic seats. This means the Republicans did take control of the House, however, speculations of a "Red Wave" did not hold up. The Democrats barely held to the Senate, with 50 seats right now, and Georgia heads to runoff. If Warnock wins as many predict, the final makeup will be 51 Democrats and 49 Republicans in the Senate. 

#Recap of Model and Prediction #

My model attempted to use aggregate data from 2018 to predict an outcome for 2022. I will go into my challenges with this later on. The general makeup of my model was that I used GLM and LM with the inputs of national economic data (including national 2018 GDP and RDI), local economic data (unemployment), weighted polling (weighted from 2018 polling data on closeness to election = more weight),  average expert prediction (1 for Strong Dem to 7 for Strong Rep), turnout, as well as the incumbency status of the winning candidate to predict the 2022 outcomes. 

As you can see, my model had many inputs and while running the regression, many of these variables didn't have much of an impact on my results. 

#Graphic Performance of My Model #
```{r old model setup, include=FALSE}
library(readr)
library(stringr)
library(dotenv)
library(jsonlite)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
library(scales)
library(rjson)
library(jtools)
library(htmlwidgets)
library(reactable)
library(sqldf)
library(blogdown)
library(ggthemes)
library(dplyr)

#Vote and seat share data from prev environments
voteseatshare_2020 <- read.csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house nationwide vote and seat share by party 1948-2020.csv")

house_party_vote_share_by_district_1948_2020 <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv")

## ECONOMIC DATA ##

# Data on GDP/quarter
economy <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/GDP_quarterly.csv")

# Data on RDI/quarter
RDI_quarterly <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/RDI_quarterly.csv")

# Data on unemployment
unemployment_national_quarterly_final <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/unemployment_national_quarterly_final.csv")

# Data on vote/seat share
popvote_df <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/house_popvote_seats.csv")

#state unemployment data
unemployment_state_monthly <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/Section data 2/unemployment_state_monthly.csv") 

#DATA FROM FEC.gov
spending_bydistrict_2018 <- read.csv("~/Desktop/Gov1347 Data/Gov1347 Projects/ElectionBlog/2018_spending_by_district.csv")

house_party_vote_share_by_district <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv")%>%
filter(raceYear == 2018) %>%
dplyr::select(State, raceYear, Area, RepCandidate, DemCandidate, RepVotesMajorPercent, DemVotesMajorPercent, st_fips, state_abb, CD, district_num, district_id, WinnerParty)

#Load data just for year 2018
house_vote_share_by_district_2018 <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv") %>%
filter(raceYear == 2018) %>%
dplyr::select(State, raceYear, Area, RepCandidate, DemCandidate, RepVotesMajorPercent, DemVotesMajorPercent, st_fips, state_abb, CD, district_num, district_id, WinnerParty)

# Change from at large states w only one district to a code-able suffix
house_vote_share_by_district_2018$CD[grep(pattern = "-AL", x = house_vote_share_by_district_2018$CD)] <- c("AK-01","DE-01","MT-01","ND-01","SD-01", "VT-01","WY-01")

#Cleaned 2018 data from Ethan
ratings_share_2018 <-read.csv("~/Desktop/2018_ratings.csv")

#Incumbent list
incumbentslist <- read.csv("~/Desktop/incumb_dist_1948-2020 (3).csv")
```

```{r W7 setup, include=FALSE}
# GROUND GAME LAB SECTION NOTES ##
dist_pv_df <- read.csv("~/Desktop/incumb_dist_1948-2020 (3).csv")
# read in cvap
cvap_district <- read.csv("~/Downloads/drive-download-20221016T234313Z-001/cvap_district_2012-2020_clean.csv")
# mutate geoid for merging
cvap_district <- cvap_district %>% dplyr::rename(st_cd_fips = geoid)

# select relevant years from voting data
table(dist_pv_df$year)
# 2012 - from 2018
# 2014, 2016, 2018, 2020 - from 2020
dist_pv_df <- dist_pv_df %>%
filter(year == 2012 | year == 2014 | year == 2018)
table(dist_pv_df$st_cd_fips)

# merge
dist_pv_cvap <- dist_pv_df %>%
inner_join(cvap_district, by = c('st_cd_fips', 'year', 'state'))

# mutate turnout
dist_pv_cvap <- dist_pv_cvap %>%
mutate(totalvotes = RepVotes + DemVotes,
turnout = totalvotes/cvap)

# mutate votes percent for glm
turnout_gen <- dist_pv_cvap %>%
mutate(DemVotesMajorPct = DemVotesMajorPercent/100,
RepVotesMajorPct = RepVotesMajorPercent/100)

# drop uncontested seats
dist_pv_cvap_closed <- dist_pv_cvap %>%
filter(!is.na(DemCandidate), !is.na(RepCandidate)) %>%
mutate(DemVotesMajorPct = DemVotesMajorPercent/100,
RepVotesMajorPct = RepVotesMajorPercent/100)

```

```{r map data setup, include=FALSE}
require(sf)

# load geographic data
get_congress_map <- function(cong=114) {
tmp_file <- tempfile()
tmp_dir <- tempdir()
zp <- sprintf("https://cdmaps.polisci.ucla.edu/shp/districts114.zip",cong)
download.file(zp, tmp_file)
unzip(zipfile = tmp_file, exdir = tmp_dir)
fpath <- paste(tmp_dir, sprintf("districtShapes/districts114.shp",cong), sep = "/")
st_read(fpath)
}

# load 114th congress 

cd114 <- get_congress_map(114)

cd114$DISTRICT <- as.numeric(cd114$DISTRICT)

cd114 <- cd114 %>% inner_join(house_party_vote_share_by_district, by= c("STATENAME"= "State", "DISTRICT" = "district_num"))


#cd114 <- cd114 %>% inner_join(local_model_data, by= c("STATENAME"= "State", "DISTRICT" = "district_num"))

districts_simp <- rmapshaper::ms_simplify(cd114, keep = 0.01)
```

```{r including turn out, include=FALSE}

turnout <- glm(turnout ~ year + state + president_party + cvap + winner_candidate_inc, data = turnout_gen)

summary(turnout)

turnoutpred_2022 <- data.frame()

turnoutpred <- as.data.frame(predict(turnout, new_data = turnoutpred_2022))

turnout_gen$prediction22 <- turnoutpred$`predict(turnout, new_data = turnoutpred_2022)`

#Added column for 2022 prediction now I can go ahead and map how well my model is prediction expected turnout v. actual turnout

#add margin column
options(scipen=999)

turnout_gen <- turnout_gen %>%
mutate(margin = turnout - prediction22)

turnout_18 <- turnout_gen %>%
filter(year == 2018)

districts_simp1 <- districts_simp %>%
dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, ENDCONG, FINALNOTE, district_id, CD, geometry)

#turnout_18 <- districts_simp1 %>% 
#left_join(turnout_18, by = 'district_id')
```


```{r local model data build, include=FALSE}
economy_q4 <- economy %>%
filter(quarter_yr == 4)

RDI_q4 <- RDI_quarterly %>%
filter(quarter_yr == 4)

unemployment_q4 <- unemployment_national_quarterly_final %>%
filter(quarter_yr == 4)

economy_model_data <- popvote_df %>%
inner_join(economy_q4, by = "year")

economy_model_data <-economy_model_data %>%
inner_join(RDI_q4, by = "year")

#Include District Level Economy Data
# Clean and update economy state level
unemployment_local <- unemployment_state_monthly %>%
filter(Year == 2018)

unemployment_local <- unemployment_local %>%
dplyr::select('State.and.area', Year, Month, Unemployed_prct)

# Clean and include local RDI data for just 2018 and then join to economy data
RDI_local <- RDI_q4 %>%
filter(year == 2018)

# RDI and unemployment merged
unemployment_local <- unemployment_local %>%
inner_join(RDI_q4, by = c("Year" = "year"))

# Include GDP here too
economy_q4 <- economy_q4 %>%
dplyr::select(year, GDP_growth_qt, GDP_growth_pct)

unemployment_local <- unemployment_local %>%
inner_join(economy_q4, by = c("Year" = "year"))

# Merge unemployment local to all state and district levels as equal. Merge by State.

unemployment_local <- unemployment_local %>% 
  dplyr::rename(State = State.and.area)


#Include District Level Polling Data (if possible)
#Using 2018 polling data to model our predictions
local_model_data_w7 <- house_vote_share_by_district_2018 %>%
inner_join(unemployment_local, by = 'State')

local_model_data_w7 <- as.data.frame(local_model_data_w7)

#local_model_data_w7 <- data_frame()

#Append the D and R Seats and national data to our local model for comparison in 2018.

natl_data_2018 <- economy_model_data %>%
dplyr::select(year, R_seats, D_seats, R_majorvote_pct, D_majorvote_pct)

local_model_data_w7 <- local_model_data_w7 %>%
inner_join(natl_data_2018, by = c("Year" = "year"))

incumbentslist18 <- incumbentslist %>%
  filter(year == 2018)

## Add expert ratings
local_model_data_w7 <- local_model_data_w7 %>%
inner_join(ratings_share_2018, by = "CD")


## Ad spend

#Adding ad spending variable to our model

#Clean up ad data and add to local model then run lm

spending_bydistrict_2018 <- spending_bydistrict_2018 %>%
dplyr::select(State, District, Candidate, Incumbent..Challenger.Open, Receipts)


#Need to add winning candidate name i think

#Only add relavent variables for incumbency
incumbentslist18 <- incumbentslist18 %>% 
  dplyr::select(state, district_id, winner_candidate_inc)


local_model_data_w7 <- local_model_data_w7 %>%
left_join(incumbentslist18, by = c("State" = "state","district_id" = "district_id"))

local_model_data_w7 <- local_model_data_w7 %>%
mutate(winner_candidate = case_when(WinnerParty == "R" ~ local_model_data_w7$RepCandidate, WinnerParty == "D" ~ local_model_data_w7$DemCandidate))


local_model_data_w7 <- local_model_data_w7 %>%
left_join(spending_bydistrict_2018, by = c("State" = "State", "district_num" ="District","winner_candidate_inc" = "Incumbent..Challenger.Open"))

#Include voting data to merge local model data and cvap information

turnoutformerge <- turnout_gen %>% 
  dplyr::select(state, year, district_num, district_id, st_cd_fips, cvap, turnout)

local_model_data_w7 <- local_model_data_w7 %>%
left_join(turnoutformerge, by = c('State' = 'state', 'district_id' = 'district_id', 'raceYear' = 'year'))

local_model_data_w7$Receipts <- as.numeric(gsub("[\\$,]", "", local_model_data_w7$Receipts))

```


```{r dataframe w7, include=FALSE}
#Take prediction for 2022 turnout and add to my existing model
w7_df <- data.frame()

#join districts and data
w7_df <-districts_simp1 %>%
left_join(local_model_data_w7, by = "district_id", "CD")

#replace missing polling values with toss up rating
w7_df$avg <- w7_df$avg %>%
replace_na(4.7)

w7_df$Unemployed_prct <- w7_df$Unemployed_prct %>%
replace_na(3.6)

w7_df$Receipts <- w7_df$Receipts %>%
replace_na(960454)

w7_df$turnout <- w7_df$turnout %>% 
  replace_na(0.515)

```


```{r w7 prediction, echo=FALSE}
w7_local_model <- glm(DemVotesMajorPercent ~ avg + Unemployed_prct + winner_candidate_inc + Receipts + turnout, data = w7_df)

summ(w7_local_model)

# add in Dem Vote prediction to w6_df and then filter for 2018, then append column to districts simp to plot
w7_df$DemVotePred22 <- predict(w7_local_model)

confint.default(w7_local_model)
```


```{r map simplify, include=FALSE}
#Simplify shapefile
w7_df <- rmapshaper::ms_simplify(w7_df, keep = 0.01)

final <- w7_df
```

```{r, read in, show_col_types = FALSE, echo=FALSE}

state_2022_data <- read_csv("~/Downloads/2022_4_0_1.csv", 
 skip = 1)

dat <- read_csv("~/Downloads/2022_4_0_3.csv")

stateFIPcode <- read.csv("~/Downloads/stateFIPcode.csv")

dat = dat[-1,]

##Add in State and then CD
dat$FIPS <- as.numeric(dat$FIPS)
dat$STATE_FIPS <- as.numeric(dat$STATE_FIPS)

##hand enter MA, ME, and MS

clean_final_22 <- final %>% 
  dplyr::select(State, district_id, st_fips, st_cd_fips, CD.x, DemVotePred22, geometry)

test <- dat %>% left_join(stateFIPcode, by = c("STATE_FIPS"= "FIPS.Code"))

test$district_id <- paste(test$Postal.Abbr., test$CD)

#Make to match district ID elsewhere
test$district_id <- str_replace_all(test$district_id, " ", "")

final_test <- test %>% 
  dplyr::select(State, district_id, FIPS, CD, `Total Vote`, Democratic, Republican, `CA Democratic`, `CA Republican`, Independent, RANKD, RANKR, RANKI)

##As of 11/21/22
final_test$FIPS <- as.character(final_test$FIPS)

##hand enter MA, ME, and MS
final_test$Democratic[final_test$FIPS == 25901] <- 153402
final_test$Republican[final_test$FIPS == 25901] <- 96499 
final_test$Democratic[final_test$FIPS == 25902] <- 178472
final_test$Republican[final_test$FIPS == 25902] <- 91100
final_test$Democratic[final_test$FIPS == 25903] <- 145507
final_test$Republican[final_test$FIPS == 25903] <- 82628
final_test$Democratic[final_test$FIPS == 25904] <- 0
final_test$Republican[final_test$FIPS == 25904] <- 0
final_test$Democratic[final_test$FIPS == 25905] <- 198617
final_test$Republican[final_test$FIPS == 25905] <- 70694
final_test$Democratic[final_test$FIPS == 25906] <- 190062
final_test$Republican[final_test$FIPS == 25906] <- 107496
final_test$Democratic[final_test$FIPS == 25907] <- 144902
final_test$Republican[final_test$FIPS == 25907] <- 26481
final_test$Democratic[final_test$FIPS == 25908] <- 184084
final_test$Republican[final_test$FIPS == 25908] <- 80961
final_test$Democratic[final_test$FIPS == 25909] <- 193426
final_test$Republican[final_test$FIPS == 25909] <- 131936

final_test$Democratic[final_test$FIPS == 23901] <- 218630
final_test$Republican[final_test$FIPS == 23901] <- 128996
final_test$Democratic[final_test$FIPS == 23902] <- 151440
final_test$Republican[final_test$FIPS == 23902] <- 140895

final_test$Democratic[final_test$FIPS == 28901] <- 45222
final_test$Republican[final_test$FIPS == 28901] <- 122122
final_test$Democratic[final_test$FIPS == 28902] <- 107071
final_test$Republican[final_test$FIPS == 28902] <- 71380
final_test$Democratic[final_test$FIPS == 28903] <- 54422
final_test$Republican[final_test$FIPS == 28903] <- 132269
final_test$Democratic[final_test$FIPS == 28904] <- 42876
final_test$Republican[final_test$FIPS == 28904] <- 127813


##LA and FL races with no reporting b/c no contest
final_test$Democratic[final_test$FIPS == 22904] <- 0
final_test$Republican[final_test$FIPS == 22904] <- 0
final_test$Democratic[final_test$FIPS == 12905] <- 0
final_test$Republican[final_test$FIPS == 12905] <- 0

final_test$Democratic <- as.numeric(final_test$Democratic)
final_test$Republican <- as.numeric(final_test$Republican)
final_test$`Total Vote`[final_test$`Total Vote` == 0] <- (final_test$Democratic + final_test$Republican)

#Alaska, Wyoming, Montana, North Dakota, South Dakota, Vermont, and Delaware <--- all at large#

clean_final_22$district_id[clean_final_22$district_id == 'AK00'] <- 'AK01'
clean_final_22$district_id[clean_final_22$district_id == 'WY00'] <- 'WY01'
clean_final_22$district_id[clean_final_22$district_id == 'MT00'] <- 'MT01'
clean_final_22$district_id[clean_final_22$district_id == 'ND00'] <- 'ND01'
clean_final_22$district_id[clean_final_22$district_id == 'SD00'] <- 'SD01'
clean_final_22$district_id[clean_final_22$district_id == 'VT00'] <- 'VT01'
clean_final_22$district_id[clean_final_22$district_id == 'DE00'] <- 'DE01'


#combine clean final with final test to make one data set for analysis 

final_22 <- final_test %>% right_join(clean_final_22, by = "district_id", geometry)

final_22 <- st_as_sf(final_22)

#Add in percentages for dem and rep for 2022

final_22$Democratic <- as.numeric(final_22$Democratic)
final_22$`Total Vote` <- as.numeric(final_22$`Total Vote`)
final_22$DemVoteActual <- (final_22$Democratic / final_22$`Total Vote`) * 100

final_22$margin22 <- final_22$DemVoteActual - final_22$DemVotePred22

```


*Below were my confidence intervals for my prediction*

```{r, add in the 2022 data into my model}

final_22 <-rmapshaper::ms_simplify(final_22, keep = 0.01)

```

```{r plot for comparison}
margin_plot <- ggplot() +
geom_sf(data= final_22, aes(fill= margin22), color = "grey60", size = 0.05) +
scale_fill_gradient2(low = "red",
mid = "white",
high = "blue",
midpoint = 0,
name = "Dem Vote Margin") +
coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 52), expand = FALSE)  +
theme(axis.title.x=element_blank(),
axis.text.x=element_blank(),
axis.ticks.x=element_blank(),
axis.title.y=element_blank(),
axis.text.y=element_blank(),
axis.ticks.y=element_blank(),
plot.title = element_text(margin = margin(0,0,10,0), hjust = 0.5)) +
labs(fill = "DemMajorVotePct Margin Difference", title = "Model's Margin Difference per District (actual - pred)")

ggplotly(margin_plot)
```


As you can see above, I have plotted the margin (Actual - My model's prediction) of the Democratic vote percent in 2022. My model greatly overestimated Democratic performance even in Blue states. There are some interesting things to note, one is how it underpredicted performance in states like South Carolina, Montana, and parts of Florida. I think this is an interesting finding when considered in conjuction with certain speculations political scientists and party leadership has had about an increase in Dem mobilization in certain Red stronghold states. 


#Where I Went Wrong #

I think the biggest mistake I made was not doing a pooled model. I think most of the reason I had such strong model predictions for Democrats is that I only used data from the year 2018. 

#If I Were to Do It Again ...#

